---
title: '`Quiz 4` Regression Models'
author: '`r if(knitr::is_html_output()) {"&#x1f468;&#x1F3FB;&#x200d;&#x1f4bb; Anderson H Uyekita"} else {NULL}`'
output:
  html_document:
    self_contained: true
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
      smooth_scroll: true
  github_document: default
params:
  author: 'Anderson H Uyekita'
  course: 'Regression Models'
  course_url: 'https://www.coursera.org/learn/regression-models'
  specialization: 'Data Science: Statistics and Machine Learning Specialization'
  specialization_url: 'https://www.coursera.org/specializations/data-science-statistics-machine-learning'
  instructor: 'Brian Caffo'
  course_start: !r lubridate::ymd('2022/07/05', tz = 'America/Sao_Paulo')
  course_finish: !r lubridate::ymd('2022/07/19', tz = 'America/Sao_Paulo')
  week: 'Week 4'
always_allow_html: true
---

```{r setup,include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, message = FALSE, warning = FALSE)
library(rmarkdown)
library(lubridate)
library(magrittr)
library(tidyverse)
library(ggplot2)
```

`r if(!knitr::is_html_output()) {sprintf(fmt = "* &#x1f468;&#x1F3FB;&#x200d;&#x1f4bb; Author: %s", params$author)}`
`r sprintf(fmt = "* &#x1f4da; Specialization: [%s](%s){target='_blank' rel='noopener'}", params$specialization, params$specialization_url)`
`r sprintf(fmt = "* &#x1f4d6; Course: [%s](%s){target='_blank' rel='noopener'}", params$course, params$course_url)`
    `r sprintf(fmt = "* &#x1F9D1;&#x200d;&#x1F3EB; Instructor: %s", params$instructor)`
`r sprintf(fmt = "* &#x1F4C6; %s", params$week)`
    `r sprintf(fmt = "* &#x1F6A6; Start: %s", format(params$course_start, "%A, %d %B %Y"))`
    `r sprintf(fmt = "* &#x1F3C1; Finish: %s", format(params$course_finish, "%A, %d %B %Y"))`
`r if(!knitr::is_html_output()) {"* &#x1F30E; Rpubs: [Interactive Document](https://rpubs.com/AndersonUyekita/quiz-4_regression-models)"}else{"* &#x1F4E6; Github Repository: [Static Document](https://github.com/AndersonUyekita/regression-models/blob/main/Week%204/quiz-4_regression-models.md){target='_blank' rel='noopener'}"}`

--------------------------------------------------------------------------------

## Question 1

Consider the space shuttle data `?shuttle` in the $MASS$ library. Consider modeling the use of the autolander as the outcome (variable name **use**). Fit a logistic regression model with autolander (variable auto) use (labeled as "auto" 1) versus not (0) as predicted by wind sign (variable **wind**). Give the estimated odds ratio for autolander use comparing head winds, labeled as "head" in the variable headwind (numerator) to tail winds (denominator).

- [ ] 1.327
- [ ] 0.031
- [x] 0.969
- [ ] -0.031

**Answer**

```{r}
# Loading MASS package to use shuttle dataset.
library(MASS)

# Exploring
str(MASS::shuttle)

# Subsetting the dataset.
MASS::shuttle %>%
    
    # Selecting only use and wind variables.
    dplyr::select(use, wind) %>%
    
    # Releveling the factors following the instructions.
    mutate(use = factor(recode(use, noauto = 0, auto = 1)),
           wind = factor(recode(wind, tail = 0, head = 1))) -> df_q1

# Fitting a Logistic Regression
fit_q1 <- glm(data = df_q1, formula = use ~ wind, family = "binomial")

# Printing the coefficients.
exp(coef(fit_q1))
```

## Question 2

Consider the previous problem. Give the estimated odds ratio for autolander use comparing head winds (numerator) to tail winds (denominator) adjusting for wind strength from the variable magn.

- [ ] 1.00
- [x] 0.969
- [ ] 0.684
- [ ] 1.485

**Answer**

```{r}
# Loading MASS package to use shuttle dataset.
library(MASS)

# Exploring
str(MASS::shuttle);

# Subsetting the dataset.
MASS::shuttle %>%
    
    # Selecting only use and wind variables.
    dplyr::select(use, wind, magn) %>%
    
    # Releveling the factors following the instructions.
    mutate(use = factor(recode(use, noauto = 0, auto = 1)),
           wind = factor(recode(wind, tail = 0, head = 1))) -> df_q2

# Fitting a Logistic Regression
fit_q2 <- glm(data = df_q2, formula = use ~ wind + magn, family = "binomial")

# Printing the summary
summary(fit_q2)$coeff;

# Printing the coefficients.
exp(coef(fit_q2))
```

## Question 3

If you fit a logistic regression model to a binary variable, for example use of the autolander, then fit a logistic regression model for one minus the outcome (not using the autolander) what happens to the coefficients?

- [x] The coefficients reverse their signs.
- [ ] The intercept changes sign, but the other coefficients don't.
- [ ] The coefficients change in a non-linear fashion.
- [ ] The coefficients get inverted (one over their previous value).

**Answer**

```{r}
# Loading MASS package to use shuttle dataset.
library(MASS)

# Exploring
str(MASS::shuttle);

# Subsetting the dataset.
MASS::shuttle %>%
    
    # Selecting only use and wind variables.
    dplyr::select(use, wind) %>%
    
    # Releveling the factors following the instructions.
    mutate(use = factor(recode(use, noauto = 1, auto = 0)),
           wind = factor(recode(wind, tail = 0, head = 1))) -> df_q3

# Fitting a Logistic Regression
fit_q3_a <- glm(data = df_q3, formula = use ~ wind, family = "binomial")
fit_q3_b <- glm(data = df_q1, formula = use ~ wind, family = "binomial")

# Printing the coefficients.
coef(fit_q3_a);coef(fit_q3_b)
```

The expression $y = 1 - x$ converts 0 into 1 and 1 into 0. So I have changed the recode function to `wind` variable. Following the instructions, I have changed the wind (not the autolander).

## Question 4

Consider the insect spray data $InsectSprays$. Fit a Poisson model using spray as a factor level. Report the estimated relative rate comapring spray A (numerator) to spray B (denominator).

- [ ] -0.056
- [ ] 0.321
- [x] 0.9457
- [ ] 0.136

**Answer**

```{r}
# Exploring Insect Spray dataset.
str(InsectSprays)

# Releveling the spray variable.
df_q4 <- InsectSprays %>%
    mutate(spray = fct_relevel(spray, c("B", "A", "C", "D", "E", "F")))

# Fitting the model.
fit_q4 <- glm(data = df_q4, formula = count ~ spray, family = "poisson")

# Printing the coefficients.
exp(coef(fit_q4))
```

The spray B is the baseline, so every spray coefficient is based on it. For this reason, the relative rate comparison between A and B is 0.9456522.


## Question 5

Consider a Poisson glm with an offset, $t$. So, for example, a model of the form $glm(count ~ x + offset(t), family = poisson)$ where $x$ is a factor variable comparing a treatment (1) to a control (0) and $t$ is the natural log of a monitoring time. What is impact of the coefficient for $x$ if we fit the model $glm(count ~ x + offset(t2), family = poisson)$ where $t2 <- log(10) + t$? In other words, what happens to the coefficients if we change the units of the offset variable.

_(Note, adding log(10) on the log scale is multiplying by 10 on the original scale.)_

- [ ] The coefficient is subtracted by log(10).
- [ ] The coefficient estimate is divided by 10.
- [x] The coefficient estimate is unchanged
- [ ] The coefficient estimate is multiplied by 10.

**Answer**

```{r}
# Selecting a dataset with treatment and control groups.
df_q5 <- datasets::PlantGrowth %>%
    filter(group %in% c("ctrl", "trt1"))

# Fitting the model - "t"
fit_q5_1 <- glm(data = df_q5, formula = weight ~ group, family = poisson, offset = log(weight))

# Fitting the model - "t2"
fit_q5_2 <- glm(data = df_q5, formula = weight ~ group, family = poisson, offset = log(weight) + log(10))

# Comparing the results.
coef(fit_q5_1);coef(fit_q5_2)
```

The slope is unchanged.


## Question 6

Consider the data

```{r}
x <- -5:5
y <- c(5.12, 3.93, 2.67, 1.87, 0.52, 0.08, 0.93, 2.05, 2.54, 3.87, 4.97)
```

Using a knot point at 0, fit a linear model that looks like a hockey stick with two lines meeting at x=0. Include an intercept term, x and the knot point term. What is the estimated slope of the line after 0?

- [ ] -1.024
- [x] 1.013
- [ ] -0.183
- [ ] 2.037

**Answer**

```{r}
# Creating the hockey stick with lines meeting in x = 0
stick <- abs(-5:5)

# Creating the data frame.
df_q6 <- data.frame(stick, x, y)


# Fitting a linear model.
fit_q6 <- lm(data = df_q6, formula = y ~ x + stick)

# Printing the coefficients.
summary(fit_q6)$coeff
```

$$y = \beta_0 + \beta_1 \cdot x + \beta_2 \cdot stick$$

$$y = -0.1825 -0.0055 \cdot x + 1.018612903 \cdot stick$$

stick = 0

$$y = -0.1825 -0.0055 \cdot x$$

stick = 1 (after the x = ZERO)

$$y = -0.1825 + 1.013067 \cdot x $$

Printing the value in a ggplot2.

```{r}
ggplot(data = df_q6, aes(x = x, y = y)) + 
    geom_point(color = "#F8766D") + 
    geom_line(aes(x = x, y = stick), color = "#619CFF") + 
    ggtitle(label = "Hockey Stick Lines in Blue", subtitle = "y is the red points")

```