---
title: '`Quiz 3 (Optional)` Regression Models'
author: '`r if(knitr::is_html_output()) {"&#x1f468;&#x1F3FB;&#x200d;&#x1f4bb; Anderson H Uyekita"} else {NULL}`'
output:
  github_document: default
  html_document:
    self_contained: true
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
      smooth_scroll: true
params:
  author: 'Anderson H Uyekita'
  course: 'Regression Models'
  course_url: 'https://www.coursera.org/learn/regression-models'
  specialization: 'Data Science: Foundations using R Specialization'
  specialization_url: 'https://www.coursera.org/specializations/data-science-foundations-r'
  instructor: 'Brian Caffo'
  course_start: !r lubridate::ymd('2022/07/05', tz = 'America/Sao_Paulo')
  course_finish: !r lubridate::today(tzone = "America/Sao_Paulo")
  week: ' Week 3'
always_allow_html: true
---

```{r setup,include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, message = FALSE, warning = FALSE)
library(rmarkdown)
library(lubridate)
```

`r if(!knitr::is_html_output()) {sprintf(fmt = "* &#x1f468;&#x1F3FB;&#x200d;&#x1f4bb; Author: %s", params$author)}`
`r sprintf(fmt = "* &#x1f4da; Specialization: [%s](%s){target='_blank' rel='noopener'}", params$specialization, params$specialization_url)`
`r sprintf(fmt = "* &#x1f4d6; Course: [%s](%s){target='_blank' rel='noopener'}", params$course, params$course_url)`
    `r sprintf(fmt = "* &#x1F9D1;&#x200d;&#x1F3EB; Instructor: %s", params$instructor)`
`r sprintf(fmt = "* &#x1F4C6; %s", params$week)`
    `r sprintf(fmt = "* &#x1F6A6; Start: %s", format(params$course_start, "%A, %d %B %Y"))`
    `r sprintf(fmt = "* &#x1F3C1; Finish: %s", format(params$course_finish, "%A, %d %B %Y"))`
`r if(!knitr::is_html_output()) {"* &#x1F30E; Rpubs: [Interactive Document](https://rpubs.com/AndersonUyekita/quiz-3_regression-models)"}else{"* &#x1F4E6; Github Repository: [Static Document](https://github.com/AndersonUyekita/regression-models/blob/main/Week%203/quiz-3_regression-models.md){target='_blank' rel='noopener'}"}`

--------------------------------------------------------------------------------

## Question 1

You are being asked to participate in a research experiment with the purpose of better understanding how people analyze data. If you complete this quiz, you are giving your consent to participate in the study. This quiz involves a short data analysis that gives you a chance to practice the regression concepts you have learned so far. 

**We anticipate that this will take about 15 minutes to complete. You will be receiving feedback on your work immediately after submission. For this reason, we ask that you do not post on the forums about this quiz to maintain the integrity of this experiment.**

> Thank you for helping us learn more about data science! -- Brian, Roger, Jeff

Your assignment is to study how income varies across different categories of college majors. You will be using data from a study of recent college graduates. Make sure to use good practices that you have learned so far in this course and previous courses in the specialization. 

If you will proceed with the analysis, click “Yes”. Otherwise, click “No” and exit the quiz.

- [x] Yes
- [ ] No

## Question 2

Your assignment is to study how income varies across college major categories. Specifically answer: “Is there an association between college major category and income?” 

To get started, start a new R/RStudio session with a clean workspace. To do this in R, you can use the q() function to quit, then reopen R. The easiest way to do this in RStudio is to quit RStudio entirely and reopen it. After you have started a new session, run the following commands. This will load a data.frame called college for you to work with.

```r
install.packages("devtools")
devtools::install_github("jhudsl/collegeIncome")
library(collegeIncome)
data(college)
```

Next download and install the matahari R package with the following commands:

```r
devtools::install_github("jhudsl/matahari")
library(matahari)
```

This package allows a record of your analysis (your R command history) to be documented. You will be uploading a file containing this record to GitHub and submitting the link as part of this quiz.

Before you start the analysis for this assignment, enter the following command to begin the documentation of your analysis:

```r
dance_start(value = FALSE, contents = FALSE)
```

You can then proceed with the rest of your analysis in R as usual. When you have finished your analysis, use the following command to save the record of your analysis on your desktop:

```r
dance_save("~/Desktop/college_major_analysis.rds")
```

Please upload this college_major_analysis.rds file to a public GitHub repository. In question 4 of this quiz, you will share the link to this file.

A codebook for the dataset is given below:

* `ank`: Rank by median earnings
* `major_code`: Major code
* `major`: Major description
* `major_category`: Category of major
* `total`: Total number of people with major
* `sample_size`: Sample size of full-time, year-round individuals used for income/earnings estimates: p25th, median, p75th
* `p25th`: 25th percentile of earnings
* `median`: Median earnings of full-time, year-round workers
* `p75th`: 75th percentile of earnings
* `perc_men`: % men with major (out of total)
* `perc_women`: % women with major (out of total)
* `perc_employed`: % employed (out of total)
* `perc_employed_fulltime`: % employed 35 hours or more (out of employed)
* `perc_employed_parttime`: % employed less than 35 hours (out of employed)
* `perc_employed_fulltime_yearround`: % employed at least 50 weeks and at least 35 hours (out of employed and full-time)
* `perc_unemployed`: % unemployed (out of employed)
* `perc_college_jobs`: % with job requiring a college degree (out of employed)
* `perc_non_college_jobs`: % with job not requiring a college degree (out of employed)
* `perc_low_wage_jobs`: % in low-wage service jobs (out of total)

**Question:** Based on your analysis, would you conclude that there is a significant association between college major category and income?

- [ ] Yes
- [ ] No

## Question 3

Please type a few sentences describing your results.

#### **Answer**

```{r,cache=TRUE,echo=TRUE}
library(collegeIncome)
data(college)
library(matahari)
```


```{r}
dance_start(value = FALSE, contents = FALSE)
```


```{r}
dance_save("~/Desktop/college_major_analysis.rds")
```



## Question 4

Please upload the file generated by matahari (college_major_analysis.rds) to a **public** GitHub repository and paste the link to that file here.



## Question 5


