---
title: '`Quiz 3` Regression Models'
author: '`r if(knitr::is_html_output()) {"&#x1f468;&#x1F3FB;&#x200d;&#x1f4bb; Anderson H Uyekita"} else {NULL}`'
output:
  html_document:
    self_contained: true
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
      smooth_scroll: true
  github_document: default
params:
  author: 'Anderson H Uyekita'
  course: 'Regression Models'
  course_url: 'https://www.coursera.org/learn/regression-models'
  specialization: 'Data Science: Statistics and Machine Learning Specialization'
  specialization_url: 'https://www.coursera.org/specializations/data-science-statistics-machine-learning'
  instructor: 'Brian Caffo'
  course_start: !r lubridate::ymd('2022/07/05', tz = 'America/Sao_Paulo')
  course_finish: !r lubridate::ymd('2022/07/18', tz = 'America/Sao_Paulo')
  week: ' Week 3'
always_allow_html: true
---

```{r setup,include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, message = FALSE, warning = FALSE)
library(rmarkdown)
library(lubridate)
```

`r if(!knitr::is_html_output()) {sprintf(fmt = "* &#x1f468;&#x1F3FB;&#x200d;&#x1f4bb; Author: %s", params$author)}`
`r sprintf(fmt = "* &#x1f4da; Specialization: [%s](%s){target='_blank' rel='noopener'}", params$specialization, params$specialization_url)`
`r sprintf(fmt = "* &#x1f4d6; Course: [%s](%s){target='_blank' rel='noopener'}", params$course, params$course_url)`
    `r sprintf(fmt = "* &#x1F9D1;&#x200d;&#x1F3EB; Instructor: %s", params$instructor)`
`r sprintf(fmt = "* &#x1F4C6; %s", params$week)`
    `r sprintf(fmt = "* &#x1F6A6; Start: %s", format(params$course_start, "%A, %d %B %Y"))`
    `r sprintf(fmt = "* &#x1F3C1; Finish: %s", format(params$course_finish, "%A, %d %B %Y"))`
`r if(!knitr::is_html_output()) {"* &#x1F30E; Rpubs: [Interactive Document](https://rpubs.com/AndersonUyekita/quiz-3_regression-models)"}else{"* &#x1F4E6; Github Repository: [Static Document](https://github.com/AndersonUyekita/regression-models/blob/main/Week%203/quiz-3_regression-models.md){target='_blank' rel='noopener'}"}`

--------------------------------------------------------------------------------

## Question 1

Consider the `mtcars` data set. Fit a model with mpg as the outcome that includes number of cylinders as a factor variable and weight as confounder. Give the adjusted estimate for the expected change in mpg comparing 8 cylinders to 4.

- [ ] 33.991
- [x] -6.071
- [ ] -3.206
- [ ] -4.256

**Answer**

```{r}
# Fitting a model using wt and cyl.
fit_q1 <- lm(data = mtcars, formula = mpg ~ wt + factor(cyl))

# Printing the coefficients.
summary(fit_q1)$coeff
```

The interpretation of `cyl` coefficients is based on cyl 4, the baseline. So the coefficient -6.07 is in comparison to the cyl 4.

## Question 2

Consider the `mtcars` data set. Fit a model with mpg as the outcome that includes number of cylinders as a factor variable and weight as a possible confounding variable. Compare the effect of 8 versus 4 cylinders on mpg for the adjusted and unadjusted by weight models.  Here, adjusted means including the weight variable as a term in the regression model and unadjusted means the model without weight included. What can be said about the effect comparing 8 and 4 cylinders after looking at models with and without weight included?

- [ ] Holding weight constant, cylinder appears to have more of an impact on mpg than if weight is disregarded.
- [x] Holding weight constant, cylinder appears to have less of an impact on mpg than if weight is disregarded. 
- [ ] Within a given weight, 8 cylinder vehicles have an expected 12 mpg drop in fuel efficiency.
- [ ] Including or excluding weight does not appear to change anything regarding the estimated impact of number of cylinders on mpg.

**Answer**

```{r}
# Printing the coefficients.
summary(lm(data = mtcars, formula = mpg ~ factor(cyl)))$coeff;
summary(lm(data = mtcars, formula = mpg ~ wt + factor(cyl)))$coeff
```

Comparing the two fitted models, when wt has disregarded the influence in cyl in greater.

## Question 3

Consider the `mtcars` data set. Fit a model with mpg as the outcome that considers number of cylinders as a factor variable and weight as confounder. Now fit a second model with mpg as the outcome model that considers the interaction between number of cylinders (as a factor variable) and weight. Give the P-value for the likelihood ratio test comparing the two models and suggest a model using 0.05 as a type I error rate significance benchmark.

- [ ] The P-value is small (less than 0.05). So, according to our criterion, we reject, which suggests that the interaction term is necessary 
- [ ] The P-value is small (less than 0.05). Thus it is surely true that there is an interaction term in the true model.
- [ ] The P-value is small (less than 0.05). So, according to our criterion, we reject, which suggests that the interaction term is not necessary. 
- [ ] The P-value is larger than 0.05. So, according to our criterion, we would fail to reject, which suggests that the interaction terms is necessary.
- [ ] The P-value is small (less than 0.05). Thus it is surely true that there is no interaction term in the true model.
- [x] The P-value is larger than 0.05. So, according to our criterion, we would fail to reject, which suggests that the interaction terms may not be necessary.

**Answer**

```{r}
fit_q3_1 <- lm(data = mtcars, formula = mpg ~ factor(cyl) + wt)
fit_q3_2 <- lm(data = mtcars, formula = mpg ~ factor(cyl) * wt)

round(summary(fit_q3_1)$coeff,4); round(summary(fit_q3_2)$coeff,4)

anova(fit_q3_1, fit_q3_2)
```

## Question 4

Consider the `mtcars` data set. Fit a model with mpg as the outcome that includes number of cylinders as a factor variable and weight inlcuded in the model as

```
lm(mpg ~ I(wt * 0.5) + factor(cyl), data = mtcars)
```

How is the `wt` coefficient interpretted?

- [ ] The estimated expected change in MPG per one ton increase in weight.
- [ ] The estimated expected change in MPG per half ton increase in weight for the average number of cylinders.
- [ ] The estimated expected change in MPG per half ton increase in weight for for a specific number of cylinders (4, 6, 8).
- [ ] The estimated expected change in MPG per half ton increase in weight.
- [x] The estimated expected change in MPG per one ton increase in weight for a specific number of cylinders (4, 6, 8).

**Answer**

```{r}
summary(lm(mpg ~ I(wt * 0.5) + factor(cyl), data = mtcars))$coeff;
summary(lm(mpg ~ wt + factor(cyl), data = mtcars))$coeff;
```

The `0.5` will not affect the interpretation or change the results of the final model.

$$mpg = \beta_0 + \beta_1 \cdot wt + \beta_2 \cdot cyl_6 + \beta_3 \cdot cyl_8$$
For model 1:

$$mpg = 33.99 -6.41 \cdot wt \cdot 0.5 -4.25 \cdot cyl_6 -6.07 \cdot cyl_8$$
For model 2:

$$mpg = 33.99 -3.206 \cdot wt -4.25 \cdot cyl_6 -6.07 \cdot cyl_8$$

In the end, both models has the same equation.

## Question 5

Consider the following data set

```{r}
x <- c(0.586, 0.166, -0.042, -0.614, 11.72)
y <- c(0.549, -0.026, -0.127, -0.751, 1.344)
```

Give the hat diagonal for the most influential point

- [ ] 0.2287
- [ ] 0.2025
- [x] 0.9946
- [ ] 0.2804

**Answer**

```{r}
max(influence(lm(y ~ x))$hat)
```

## Question 6

Consider the following data set

```{r}
x <- c(0.586, 0.166, -0.042, -0.614, 11.72)
y <- c(0.549, -0.026, -0.127, -0.751, 1.344)
```

Give the slope dfbeta for the point with the highest hat value.

- [x] -134
- [ ] 0.673
- [ ] -.00134
- [ ] -0.378

**Answer**

```{r}
influence.measures(lm(y ~ x))
```

## Question 7

Consider a regression relationship between Y and X with and without adjustment for a third variable Z. Which of the following is true about comparing the regression coefficient between Y and X with and without adjustment for Z.

- [ ] For the the coefficient to change sign, there must be a significant interaction term.
- [x] It is possible for the coefficient to reverse sign after adjustment. For example, it can be strongly significant and positive before adjustment and strongly significant and negative after adjustment.
- [ ] Adjusting for another variable can only attenuate the coefficient toward zero. It can't materially change sign.
- [ ] The coefficient can't change sign after adjustment, except for slight numerical pathological cases. 
