---
title: '`Quiz 1` Regression Models'
author: '`r if(knitr::is_html_output()) {"&#x1f468;&#x1F3FB;&#x200d;&#x1f4bb; Anderson H Uyekita"} else {NULL}`'
output:
  github_document: default
  html_document:
    self_contained: true
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
      smooth_scroll: true
params:
  author: 'Anderson H Uyekita'
  course: 'Regression Models'
  course_url: 'https://www.coursera.org/learn/regression-models'
  specialization: 'Data Science: Foundations using R Specialization'
  specialization_url: 'https://www.coursera.org/specializations/data-science-foundations-r'
  instructor: 'Brian Caffo'
  course_start: !r lubridate::ymd('2022/07/05', tz = 'America/Sao_Paulo')
  course_finish: !r lubridate::ymd('2022/07/18', tz = 'America/Sao_Paulo')
  week: ' Week 1'
always_allow_html: true
---

```{r setup,include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, message = FALSE, warning = FALSE)
library(rmarkdown)
library(lubridate)
```

`r if(!knitr::is_html_output()) {sprintf(fmt = "* &#x1f468;&#x1F3FB;&#x200d;&#x1f4bb; Author: %s", params$author)}`
`r sprintf(fmt = "* &#x1f4da; Specialization: [%s](%s){target='_blank' rel='noopener'}", params$specialization, params$specialization_url)`
`r sprintf(fmt = "* &#x1f4d6; Course: [%s](%s){target='_blank' rel='noopener'}", params$course, params$course_url)`
    `r sprintf(fmt = "* &#x1F9D1;&#x200d;&#x1F3EB; Instructor: %s", params$instructor)`
`r sprintf(fmt = "* &#x1F4C6; %s", params$week)`
    `r sprintf(fmt = "* &#x1F6A6; Start: %s", format(params$course_start, "%A, %d %B %Y"))`
    `r sprintf(fmt = "* &#x1F3C1; Finish: %s", format(params$course_finish, "%A, %d %B %Y"))`
`r if(!knitr::is_html_output()) {"* &#x1F30E; Rpubs: [Interactive Document](https://rpubs.com/AndersonUyekita/quiz-1_regression-models)"}else{"* &#x1F4E6; Github Repository: [Static Document](https://github.com/AndersonUyekita/regression-models/blob/main/Week%201/quiz-1_regression-models.md){target='_blank' rel='noopener'}"}`

--------------------------------------------------------------------------------

## Question 1
Consider the data set given below

```{r}
x <- c(0.18, -1.54, 0.42, 0.95)
```

And weights given by

```{r}
w <- c(2, 1, 3, 1)
```

Give the value of $\mu$ that minimizes the least squares equation

$$\sum_{i = 1}^{n}{w_i \cdot (x_i - \mu)^2}$$

- [ ] 0.300
- [x] 0.1471
- [ ] 1.077
- [ ] 0.0025

**Answer**

The $\mu$ value that minimizes the given function should be the mean. For this reason, I will calculate the weighted average.

$$\mu = \frac{\sum{x_i \cdot w_i}}{\sum{w_i}}$$
```{r}
# Calculating the weighted average.
sum(x * w)/sum(w)
```

## Question 2

Consider the following data set:

```{r}
x <- c(0.8, 0.47, 0.51, 0.73, 0.36, 0.58, 0.57, 0.85, 0.44, 0.42)
y <- c(1.39, 0.72, 1.55, 0.48, 1.19, -1.59, 1.23, -0.65, 1.49, 0.05)
```
Fit the regression through the origin and get the slope treating y as the outcome and x as the regressor.

_(Hint, do not center the data since we want regression through the origin, not through the means of the data.)_

- [ ] 0.59915
- [ ] -0.04462
- [x] 0.8263
- [ ] -1.713

**Answer**

Based on the `lm()` function, it is necessary to set the regression without Interceptor.

```{r}
# Creating a data frame.
df_q2 <- data.frame(x, y)

# Fitting a linear regression model without interceptor (using -1).
fit_q2 <- lm(data = df_q2, formula = y ~ x - 1)

# Printing the coefficients
summary(fit_q2)$coeff
```

## Question 3

Do `data(mtcars)` from the datasets package and fit the regression model with **mpg** as the outcome and **weight** as the predictor. Give the slope coefficient.

- [ ] 30.2851
- [ ] 0.5591
- [ ] -9.559
- [x] -5.344

**Answer**

The model is quite simple:

```{r}
# Fitting a model using mpg and weight.
fit_q3 <- lm(data = mtcars, formula = mpg ~ wt)

# Printing the coefficients
summary(fit_q3)$coeff
```

## Question 4

Consider data with an outcome (Y) and a predictor (X). The standard deviation of the predictor is one half that of the outcome. The correlation between the two variables is `0.5`. What value would the slope coefficient for the regression model with $Y$ as the outcome and $X$ as the predictor?

- [ ] 3
- [x] 1
- [ ] 4
- [ ] 0.25

**Answer**

From the given formula (1) to calculate the $\beta_1$.

$$\beta_1 = \frac{Cov(Y,X)}{(sd(X))^2}$$
The Correlation is given by the follow formula (2):

$$Cor(Y,X) = \frac{Cov(Y,X)}{sd(X) \cdot sd(Y)}$$

Using the formula (2) in (1):

$$\beta_1 = Cor(X,Y) \cdot \frac{sd(Y)}{sd(X)}$$

```{r}
# Creating the variables
cor_x_y <- 0.5
sdy_sdx <- 2

# Calculating the beta_1
beta_1 <- cor_x_y * sdy_sdx

# Printing beta_1
beta_1
```


## Question 5

Students were given two hard tests and scores were normalized to have empirical mean 0 and variance 1. The correlation between the scores on the two tests was 0.4. What would be the expected score on Quiz 2 for a student who had a normalized score of 1.5 on Quiz 1?

- [ ] 0.16
- [ ] 0.4
- [x] 0.6
- [ ] 1.0

**Answer**

In this case of normalized data, the slope will equal the correlation. Due to the mean equal to zero, the intercept should be zero.

$$y = \underbrace{\beta_0}_{\text{Should be zero}} + \underbrace{\beta_1}_{\text{Should be equal to Cor(X,Y)}} \cdot x$$
```{r}
# Creating the regression as a function.
q6 <- function(x) {
    
    beta_0 <- 0
    beta_1 <- 0.4
    
    return(beta_0 + beta_1 * x)
    }

# Calculating the value to 1.5.
q6(1.5)
```


## Question 6

Consider the data given by the following

```{r}
x <- c(8.58, 10.46, 9.01, 9.64, 8.86)
```

What is the value of the first measurement if x were normalized (to have mean 0 and variance 1)?

- [ ] 8.86
- [x] -0.9719
- [ ] 8.58
- [ ] 9.31

**Answer**

```{r}
# Normalizing th x vector.
x_norm <- (x - mean(x))/sd(x)

# Printing the first element
x_norm[1]
```

## Question 7

Consider the following data set (used above as well). What is the intercept for fitting the model with x as the predictor and y as the outcome?

```{r}
x <- c(0.8, 0.47, 0.51, 0.73, 0.36, 0.58, 0.57, 0.85, 0.44, 0.42)
y <- c(1.39, 0.72, 1.55, 0.48, 1.19, -1.59, 1.23, -0.65, 1.49, 0.05)
```

- [ ] 2.105
- [x] 1.567
- [ ] 1.252
- [ ] -1.713

**Answer**

```{r}
# Creating a data frame.
df_q7 <- data.frame(x,y)


# Fitting a model with intercept
fit_q7 <- lm(data = df_q7, formula = y ~ x)

# Printing the coefficients
summary(fit_q7)$coeff
```

## Question 8

You know that both the predictor and  response have mean 0. What can be said about the intercept when you fit a linear regression?

- [ ] Nothing about the intercept can be said from the information given.
- [x] It must be identically 0.
- [ ] It is undefined as you have to divide by zero.
- [ ] It must be exactly one.

**Answer**

```{r}
# Using the same values of Question 7
x <- c(0.8, 0.47, 0.51, 0.73, 0.36, 0.58, 0.57, 0.85, 0.44, 0.42)
y <- c(1.39, 0.72, 1.55, 0.48, 1.19, -1.59, 1.23, -0.65, 1.49, 0.05)

# Centering the data to have mean zero.
x_c <- x - mean(x)
y_c <- y - mean(y)

# Creating a data frame
df_q8 <- data.frame(x_c, y_c)

# Fitting a model
fit_q8 <- lm(data = df_q8, formula = y_c ~ x_c)

# Printing the coefficients
summary(fit_q8)$coeff
```

## Question 9

Consider the data given by

```{r}
x <- c(0.8, 0.47, 0.51, 0.73, 0.36, 0.58, 0.57, 0.85, 0.44, 0.42)
```

What value minimizes the sum of the squared distances between these points and itself?

- [ ] 0.36
- [ ] 0.8
- [x] 0.573
- [ ] 0.44

**Answer**

The answer is similar to question 1, except this data has no weight. So, the value which minimizes the Squared Distances should be the mean of x.

```{r}
# Calculating the average.
mean(x)
```

## Question 10

Let the slope having fit Y as the outcome and X as the predictor be denoted as $\beta_1$. Let the slope from fitting X as the outcome and Y as the predictor be denoted as $\gamma_1$. Suppose that you divide $\beta_1$ by $\gamma_1$; in other words consider  $\frac{\beta_1}{\gamma_1}$. What is this ratio always equal to?

- [x] $Var(Y)/Var(X)$
- [ ] $2 \cdot sd(X)/sd(Y)$
- [ ] 1
- [ ] $Cor(Y,X)$

**Answer**

Given $\beta_1$:

$$\beta_1 = Cor(Y,X) \cdot \frac{sd(Y)}{sd(X)}$$

Given $\gamma_1$:

$$\gamma_1 = Cor(Y,X) \cdot \frac{sd(Y)}{sd(X)}$$

Calculating the $\frac{\beta_1}{\gamma_1}$:

$$\frac{\beta_1}{\gamma_1} = \frac{Cor(Y,X) \cdot \frac{sd(Y)}{sd(X)}}{Cor(Y,X) \cdot \frac{sd(Y)}{sd(X)}}=\Big( \frac{sd(Y)}{sd(X)} \Big)^2 = \frac{Var(Y)}{Var(X)}$$
